An Analysis of Homelessness in Toronto in regards to COVID-19 and other factors

### 2022-02-07 outputs/paper/paper.Rmd
Finished graphs and tables

---
title: "Accessment of Homelessness in Toronto in regards to COVID-19 and other factors"
subtitle: "Data from January 2020 to December 2021"
author: "Ziyao Han"
thanks: "Code and data are available at: https://github.com/HanFrank/STA304-Paper-1"
date: "February 6, 2022"
abstract: "First sentence. Second sentence. Third sentence. Fourth sentence."
output: 
  bookdown::pdf_document2
toc: FALSE
bibliography: references.bib
---

@@

First Paragraph is going to be motivational and broad.

@@

Third paragraph is about implications.

Final Paragraph about the remainder of the paper.

@@

## Covid-19 and its effects on homelessness shelter {#covid}

```{r graph1, fig.cap="Active number of homeless over time", echo=FALSE, fig.height=4, fig.width=6}
# Creating subset of only total population
population_plot <- monthly_shelter_usage[monthly_shelter_usage$population_group == "All Population",]
# Plotting total usage of homeless shelter over time and overlaying start of covid
population_plot %>%
  ggplot(aes(x = date, y = actively_homeless)) +
  labs(x = "Date", y = "Actively Homeless", title =  "Usage of Homeless shelters dropping due to Covid-19") +
  geom_line() +
  geom_line(aes(x=as.Date("2020-03-01")), linetype="dotted", col = "red") + # Drawing dotted red line for covid
  theme_minimal()
```

@@

## Age demographics

## Areas of further exploration

>>>>

---
title: "An Analysis of Homelessness in Toronto in regards to COVID-19 and other factors"
subtitle: "Data from January 2020 to December 2021"
author: "Ziyao Han"
thanks: "Code and data are available at: https://github.com/HanFrank/STA304-Paper-1"
date: "February 6, 2022"
abstract: "First sentence. Second sentence. Third sentence. Fourth sentence."
output: 
  bookdown::pdf_document2:
    extra_dependencies: ["float"]
toc: TRUE
bibliography: references.bib
---

@@

First Paragraph is going to be motivational and broad.
importance of homelessness shelters for a city

@@

Third paragraph is about implications.
covid19 has changed the way the world works
- social distancing
- mask requirements
- vaccines

@@

## Covid-19 and its effects on homelessness shelter usage {#covid}

```{r graph1, fig.cap="Active number of homeless over time", echo=FALSE, fig.height=4, fig.width=6}
# Creating subset of only total population
total_population <- monthly_shelter_usage[monthly_shelter_usage$population_group == "All Population",]
# Plotting total usage of homeless shelter over time and overlaying start of covid
ggplot(total_population, aes(x = date, y = actively_homeless)) +
geom_line() +
# Drawing dotted red line for covid
geom_line(aes(x=as.Date("2020-03-01"), col = "COVID-19 response begins in Toronto"), linetype="dotted") + 
labs(x = "Date", 
     y = "Actively Homeless", 
     title =  "Usage of Homeless shelters dropping due to COVID-19",
     col = "Event: ") +
theme_minimal() +
theme(legend.position = "bottom")
```

Stuff about this graph

### Chronic shelter usage is stable

```{r graph2, fig.cap="Active number of homeless over time for each population group", echo=FALSE}
ggplot(monthly_shelter_usage, mapping = aes(x = date, y = actively_homeless, colour = population_group)) +
geom_line() +
labs(x = "Date", 
     y = "Actively Homeless", 
     title =  "Chronic shelter usage remains stable through COVID-19",
     caption = "Began tracking Indigenous numbers in Jan 2021",) +
theme_minimal()
```

@@

```{r gendertable, echo = FALSE, fig.cap="Active Homeless by Gender"}
# Creating empty tibble for gender
gender_table <- tibble("Gender" = character(),
                       "Total" = numeric(),
                       "Percentage" = numeric())

# Populating dataframe
gender_table <- add_row(gender_table, 
                        Gender = "Male", 
                        Total = sum(total_population$gender_male), 
                        Percentage = round(sum(total_population$gender_male) / sum(total_population$actively_homeless) * 100, digits = 1))

gender_table <- add_row(gender_table, 
                        Gender = "Female", 
                        Total = sum(total_population$gender_female),
                        Percentage = round(sum(total_population$gender_female) / sum(total_population$actively_homeless) * 100, digits = 1))

gender_table <- add_row(gender_table,
                        Gender = "Transgender, Non-binary or Two-spirit", 
                        Total = sum(total_population$`gender_transgender,non-binary_or_two_spirit`), 
                        Percentage = round(sum(total_population$`gender_transgender,non-binary_or_two_spirit`) / sum(total_population$actively_homeless) * 100, digits = 1))

# Creating kable
gender_table %>% 
  kable(
    caption = "Table of shelter usage by gender",
    digits = 1,
    booktabs = TRUE,
    linesep=" ",
  ) %>% 
    add_footnote("Percentage ± 0.1", notation = "number") # Adding rounding error
```
## Movement to Permanant

```{r, fig.cap="Movement between Permanant housing", echo=FALSE}
ggplot(total_population, aes(x = date)) +
geom_line(aes(y = moved_to_housing, col = "Moved")) +
geom_line(aes(y = returned_from_housing, col = "Returned")) + 
# Drawing dotted red line for covid
labs(x = "Date", 
     y = "# of people", 
     title =  "People moving to permanant housing dropping significantly",
     col = "Action: ") +
theme_minimal() +
theme(legend.position = "bottom")
```

## Areas of potential further exploration
